How fluid-mechanical erosion creates anisotropic porous media

### 2022-06-22 docs/PhysicsPaper/StokesErosion.tex
NM

%-----------------------------------------------------------------------------------------------%
\section{Boundary Integral Equation and Cauchy Integral Formulation}
\label{sec:DLP}
%-----------------------------------------------------------------------------------------------%
To accurately and efficiently solve the Stokes
  equations~\eqref{eqn:StokesEq}, we reformulate the system as a
  boundary integral equation (BIE). A BIE formulation has the advantage
  that all the unknowns are on the one-dimensional boundary of the
  domain. Therefore, only the boundary of the complex geometry must be
  discretized which we do with a spectrally accurate Fourier basis.
  Applying the same approach as our previous
  works~\citep{quaife2018boundary, chiu2020viscous}, we represent the
  velocity as the sum of a double-layer potential and a combination of
  Stokeslets and rotlets~\cite{pow-mir1987}
\begin{align}
  \uu(\xx) = \DDD[\eeta](\xx) + \sum_{\ell=1}^{M}\left(
    S[\llambda_\ell](\xx) + R[\xi_\ell](\xx)\right), 
    \quad \xx \in \Omega,
  \label{eqn:velocity}
\end{align}
where 
\begin{align}
  \DDD[\eeta](\xx) = \frac{1}{\pi}\int_{\bd\Omega} 
  \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2}
  \eeta(\yy) \, ds_\yy.
  \label{eqn:velocityDLP}
\end{align}
Note that $\bd\Omega$ includes both the eroding bodies and the outer
boundary. The Stokeslet and the rotlets are
\begin{align}
  S[\lambda_\ell](\xx) = \frac{1}{4\pi} \left(-\log \rho_\ell +
  \frac{\rr_\ell \otimes \rr_\ell}{\rho_\ell^2} \right) \llambda_\ell,
  \quad
  R[\xi_\ell](\xx) = \frac{\rr_\ell^\perp}{\rho_\ell^2}\xi_\ell,
\end{align}
respectively, where $\rr_\ell = \xx - \cc_\ell$, $\rho_\ell =
\|\rr_\ell\|$, and $\cc_\ell$ is a point inside body $\ell$. If the
density function, Stokeslets, and rotlets satisfy the second-kind
boundary integral equation
\begin{subequations}
  \label{eqn:FredBIE}
  \begin{alignat}{3}
    \UU(\xx) &= -\frac{1}{2}\eeta(\xx) + \DDD[\eeta](\xx) + 
      \sum_{\ell=1}^{M}\left(S[\llambda_\ell](\xx) + R[\xi_\ell](\xx)\right), 
      \quad &&\xx \in \Gamma, \\
    \mathbf{0} &= -\frac{1}{2}\eeta(\xx) + \DDD[\eeta](\xx) + 
      \sum_{\ell=1}^{M}\left(S[\llambda_\ell](\xx) + R[\xi_\ell](\xx)\right), 
      &&\xx \in \gamma_\ell, \: \ell = 1,\ldots,M, \\
    \llambda_\ell &= \int_{\gamma_\ell} \eeta(\yy) \, ds_\yy, 
      &&\ell = 1,\ldots,M, \\
    \xi_\ell &= \int_{\gamma_\ell} (\yy - \cc_\ell)^\perp \cdot 
      \eeta(\yy) \, ds_\yy, &&\ell = 1,\ldots,M,
  \end{alignat}
\end{subequations}
then the representation~\eqref{eqn:velocity} satisfies the Stokes
equations with the required boundary conditions~\eqref{eqn:StokesEq}. We
solve~\eqref{eqn:FredBIE} by discretizing $\gamma_\ell$ and $\Gamma$ at
equispaced collocation points, and then replacing the integrals with
quadrature rules. This results in a linear system with a
mesh-independent condition number, and it is solved iteratively with
GMRES.

Instead of evaluating the Stokes double-layer
potential~\eqref{eqn:velocityDLP} as a contour integral in $\RR^2$, we
convert the integral to a sum of contour integrals around Jordan curves
in $\CC$. We identify $\xx = (x_1,x_2)$ with $x = x_1 + ix_x$, and use
similar notation for $\yy$, $\nn$, and $\eeta$. We also interpret
$\gamma$, the boundary of the $i^{th}$ grain or the bounding box
$\Gamma$, as a Jordan curve in $\CC$. We introduce the functions
\begin{align}
  \tau_1(y) = \eta(y) \overline{n(y)} \Re(n(y)), \quad 
  \tau_2(y) = \eta(y) \overline{n(y)} \Im(n(y)).
\end{align}
We then define the five Cauchy integrals
\begin{align}
  v_1(x) &= \frac{1}{2\pi i} \int_{\gamma} \frac{Re(\eta(y))}{x-y} \,dy, 
  \quad
  v_2(x) = \frac{1}{2\pi i} \int_{\gamma} \frac{Im(\eta(y))}{x-y} \, dy, 
  \quad
  v_3(x) = \frac{1}{2\pi i} \int_{\gamma} \frac{Re(\overline{y}\eta(y))}{x-y} \, dy, \\
  v_4(x) &= \frac{1}{2\pi i} \int_{\gamma} \frac{\tau_1(y)}{x-y} \, dy,
  \quad
  v_5(x) = \frac{1}{2\pi i} \int_{\gamma} \frac{\tau_2(y)}{x-y} \, dy.
\end{align}
Then, the first and second components of the Stokes double-layer
potential are
\begin{subequations}
  \begin{align}
    u_1(x) &= -\Re(x)\Re(v_1'(x)) - \Im(x)\Re(v_2'(x)) + 
             \Re(v_3'(x)) + \Re(v_4(x)) \\
    u_2(x) &= +\Re(x)\Im(v_1'(x)) + \Im(x)\Im(v_2'(x)) - 
             \Im(v_3'(x)) + \Re(v_5(x)),
  \end{align}
  \label{eqn:velocityCauchy}
\end{subequations}
respectively. The components of the deformation tensor are
\begin{subequations}
  \begin{align}
    \pderiv{u_1}{x_1} &= -\Re(v'_1(x)) + \Re(v''_3(x)) + \Re(v'_4(x))
                         -\Re(x)\Re(v''_1(x)) - \Im(x)\Re(v''_2(x)) \\
    \pderiv{u_1}{x_2} &= -\Re(v'_2(x)) - \Im(v''_3(x)) - \Im(v'_4(x))
                         +\Re(x)\Im(v''_1(x)) + \Im(x)\Im(v''_2(x)) \\
    \pderiv{u_2}{x_1} &= +\Im(v'_1(x)) - \Im(v''_3(x)) + \Re(v'_5(x))
                         +\Re(x)\Im(v''_1(x)) + \Im(x)\Im(v''_2(x)) \\
    \pderiv{u_2}{x_2} &= +\Im(v'_2(x)) - \Re(v''_3(x)) - \Im(v'_5(x))
                         +\Re(x)\Re(v''_1(x)) + \Im(x)\Re(v''_2(x)) 
  \end{align}
  \label{eqn:deformationCauchy}
\end{subequations}
Having the deformation tensor at hand, the vorticity can be shown to
satisfy
\begin{align}
  \omega(x) = \pderiv{u_2}{x_1} - \pderiv{u_1}{x_2} = 
     \Im(v_1'(x)) + \Re(v_2'(x)) + \Re(v_5'(x)) + \Im(v_4'(x)).
  \label{eqn:vorticityCauchy}
\end{align}
We note that the deformation tensor requires second-order derivatives of
Cauchy integrals, while the vorticity only requires first-order
derivatives.

@@

With the flow and shear stresses computed, we seek to evolve the boundaries of erodable bodies. For numerical stability, we modify \eqref{Vn0} to include a smoothing term that depends on local curvature $\kappa$, as in previous work \citep{quaife2018boundary}
\begin{align}
  \Vn = \CE \, \abs{\tau} + \epsilon \langle\abs{\tau}\rangle \left(
    \frac{L}{2\pi} \kappa - 1 \right),
\end{align}
The last term is a smoothing term with strength $\epsilon \ll 1$ and proportional to the spatial average of the shear stress $\langle\abs{\tau}\rangle$. $L$ indicates the total arc length of the body, and, inside the parenthesis,  the mean curvature is subtracted so that this term preserves area. As such, the only source of material loss is the first term, $\CE \, \abs{\tau}$, which represents the shear-dependent erosion law. In addition, we apply a narrow Gaussian filter to the distribution $\abs{\tau}$ to further improve stability.

>>>>

%-----------------------------------------------------------------------------------------------%
\section{Boundary Integral Equation and Cauchy Integral Formulation}
\label{sec:DLP}
%-----------------------------------------------------------------------------------------------%
To accurately and efficiently solve the Stokes
  equations~\eqref{eqn:StokesEq}, we reformulate the system as a
  boundary integral equation (BIE). A BIE formulation has the advantage
  that all the unknowns are on the one-dimensional boundary of the
  domain. Therefore, only the boundary of the complex geometry must be
  discretized which we do with a spectrally accurate Fourier basis.
  Applying the same approach as our previous
  works~\citep{quaife2018boundary, chiu2020viscous}, we represent the
  velocity as the sum of a double-layer potential and a combination of
  Stokeslets and rotlets~\cite{pow-mir1987}
\begin{align}
  \uu(\xx) = \DDD[\eeta](\xx) + \sum_{\ell=1}^{M}\left(
    S[\llambda_\ell](\xx) + R[\xi_\ell](\xx)\right), 
    \quad \xx \in \Omega,
  \label{eqn:velocity}
\end{align}
where 
\begin{align}
  \DDD[\eeta](\xx) = \frac{1}{\pi}\int_{\bd\Omega} 
  \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2}
  \eeta(\yy) \, ds_\yy.
  \label{eqn:velocityDLP}
\end{align}
\todo{Maybe define what $\rr$ is.}
Note that $\bd\Omega$ includes both the eroding bodies and the outer
boundary. The Stokeslet and the rotlets are
\begin{align}
  S[\lambda_\ell](\xx) = \frac{1}{4\pi} \left(-\log \rho_\ell +
  \frac{\rr_\ell \otimes \rr_\ell}{\rho_\ell^2} \right) \llambda_\ell,
  \quad
  R[\xi_\ell](\xx) = \frac{\rr_\ell^\perp}{\rho_\ell^2}\xi_\ell,
\end{align}
respectively, where $\rr_\ell = \xx - \cc_\ell$, $\rho_\ell =
\|\rr_\ell\|$, and $\cc_\ell$ is a point inside body $\ell$. If the
density function, Stokeslets, and rotlets satisfy the second-kind
boundary integral equation
\begin{subequations}
  \label{eqn:FredBIE}
  \begin{alignat}{3}
    \UU(\xx) &= -\frac{1}{2}\eeta(\xx) + \DDD[\eeta](\xx) + 
      \sum_{\ell=1}^{M}\left(S[\llambda_\ell](\xx) + R[\xi_\ell](\xx)\right), 
      \quad &&\xx \in \Gamma, \\
    \mathbf{0} &= -\frac{1}{2}\eeta(\xx) + \DDD[\eeta](\xx) + 
      \sum_{\ell=1}^{M}\left(S[\llambda_\ell](\xx) + R[\xi_\ell](\xx)\right), 
      &&\xx \in \gamma_\ell, \: \ell = 1,\ldots,M, \\
    \llambda_\ell &= \int_{\gamma_\ell} \eeta(\yy) \, ds_\yy, 
      &&\ell = 1,\ldots,M, \\
    \xi_\ell &= \int_{\gamma_\ell} (\yy - \cc_\ell)^\perp \cdot 
      \eeta(\yy) \, ds_\yy, &&\ell = 1,\ldots,M,
  \end{alignat}
\end{subequations}
then the representation~\eqref{eqn:velocity} satisfies the Stokes
equations with the required boundary conditions~\eqref{eqn:StokesEq}. We
solve~\eqref{eqn:FredBIE} by discretizing $\gamma_\ell$ and $\Gamma$ at
equispaced collocation points, and then replacing the integrals with
quadrature rules. This results in a linear system with a
mesh-independent condition number, and it is solved iteratively with
GMRES.

Instead of evaluating the Stokes double-layer
potential~\eqref{eqn:velocityDLP} as a contour integral in $\RR^2$, we
convert the integral to a sum of contour integrals around Jordan curves
in $\CC$. We identify $\xx = (x_1,x_2)$ with $x = x_1 + ix_x$ \todo{It looks like there is a typo with $x_x$.}
and use
similar notation for $\yy$, $\nn$, and $\eeta$. We also interpret
$\gamma$, the boundary of the $i^{th}$ grain or the bounding box
$\Gamma$, as a Jordan curve in $\CC$. We introduce the functions
\begin{align}
  \tau_1(y) = \eta(y) \overline{n(y)} \Re(n(y)), \quad 
  \tau_2(y) = \eta(y) \overline{n(y)} \Im(n(y)).
\end{align}
We then define the five Cauchy integrals
\begin{align}
  v_1(x) &= \frac{1}{2\pi i} \int_{\gamma} \frac{Re(\eta(y))}{x-y} \,dy, 
  \quad
  v_2(x) = \frac{1}{2\pi i} \int_{\gamma} \frac{Im(\eta(y))}{x-y} \, dy, 
  \quad
  v_3(x) = \frac{1}{2\pi i} \int_{\gamma} \frac{Re(\overline{y}\eta(y))}{x-y} \, dy, \\
  v_4(x) &= \frac{1}{2\pi i} \int_{\gamma} \frac{\tau_1(y)}{x-y} \, dy,
  \quad
  v_5(x) = \frac{1}{2\pi i} \int_{\gamma} \frac{\tau_2(y)}{x-y} \, dy.
\end{align}
Then, the first and second components of the Stokes double-layer
potential are
\begin{subequations}
  \begin{align}
    u_1(x) &= -\Re(x)\Re(v_1'(x)) - \Im(x)\Re(v_2'(x)) + 
             \Re(v_3'(x)) + \Re(v_4(x)) \\
    u_2(x) &= +\Re(x)\Im(v_1'(x)) + \Im(x)\Im(v_2'(x)) - 
             \Im(v_3'(x)) + \Re(v_5(x)),
  \end{align}
  \label{eqn:velocityCauchy}
\end{subequations}
respectively. The components of the deformation tensor are
\begin{subequations}
  \begin{align}
    \pderiv{u_1}{x_1} &= -\Re(v'_1(x)) + \Re(v''_3(x)) + \Re(v'_4(x))
                         -\Re(x)\Re(v''_1(x)) - \Im(x)\Re(v''_2(x)) \\
    \pderiv{u_1}{x_2} &= -\Re(v'_2(x)) - \Im(v''_3(x)) - \Im(v'_4(x))
                         +\Re(x)\Im(v''_1(x)) + \Im(x)\Im(v''_2(x)) \\
    \pderiv{u_2}{x_1} &= +\Im(v'_1(x)) - \Im(v''_3(x)) + \Re(v'_5(x))
                         +\Re(x)\Im(v''_1(x)) + \Im(x)\Im(v''_2(x)) \\
    \pderiv{u_2}{x_2} &= +\Im(v'_2(x)) - \Re(v''_3(x)) - \Im(v'_5(x))
                         +\Re(x)\Re(v''_1(x)) + \Im(x)\Re(v''_2(x)) 
  \end{align}
  \label{eqn:deformationCauchy}
\end{subequations}
Having the deformation tensor at hand, the vorticity can be shown to
satisfy
\begin{align}
  \omega(x) = \pderiv{u_2}{x_1} - \pderiv{u_1}{x_2} = 
     \Im(v_1'(x)) + \Re(v_2'(x)) + \Re(v_5'(x)) + \Im(v_4'(x)).
  \label{eqn:vorticityCauchy}
\end{align}
We note that the deformation tensor requires second-order derivatives of
Cauchy integrals, while the vorticity only requires first-order
derivatives.

@@

With the flow and shear stresses computed, we seek to evolve the boundaries of erodable bodies. For numerical stability, we modify \eqref{Vn0} to include a smoothing term that depends on local curvature $\kappa$, as in previous work \citep{quaife2018boundary}
\begin{align}
  \Vn = \CE \, \abs{\tau} + \epsilon \langle\abs{\tau}\rangle \left(
    \frac{L}{2\pi} \kappa - 1 \right),
\end{align}
The last term is a smoothing term that has strength $\epsilon \ll 1$ and that scales with spatial average of the shear stress $\langle\abs{\tau}\rangle$. $L$ indicates the total arc length of the body, and, inside the parenthesis,  the mean curvature is subtracted so that this term preserves area. As such, the only source of material loss is the first term, $\CE \, \abs{\tau}$, which represents the shear-dependent erosion law. In addition, we apply a narrow Gaussian filter to the distribution $\abs{\tau}$ to further improve stability.
